### 2013-07-02 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
updating figure locations in parallel chapter

Compared to the serial construction of the Neumann-Ulam decomposition,
data for all states that is required to be on process must be
collected in parallel. For the pure domain decomposition case, given
an input parallel decomposition for the Neumann-Ulam decomposition,
for all local states $m$ we require access to $P_{mn}$ and $W_{mn}$
(or $P^T_{mn}$ and $W^T_{mn}$) for the adjoint method for all possible
states $n$. Given by Figure~\ref{fig:diffusion_graph}, consider the
adjacency graph of a single state in the neutron diffusion matrix for
the model problem presented in Appendix~\ref{chap:diffusion_problem}.
\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{ \input{chapters/parallel_mc/stencil_graph.pdftex_t} }
  \end{center}
  \caption{\textbf{Neutron diffusion equation state adjacency graph.}
    \textit{Adjacent boundary states are collected by traversing the
      graph one step for all local boundary states and collecting the
      equations for those states that are not local. For a given
      transition in the random walk for the diffusion problem, a
      history at mesh point $(i,j)$ may transition to all adjacent
      mesh points including itself, corresponding to a global state
      transition from $m$ to $n$ in the linear system.}}
  \label{fig:diffusion_graph}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.5}{ \input{chapters/parallel_mc/domain_to_domain.pdftex_t} }
  \end{center}
  \caption{\textbf{Nearest neighbor history communication sequence.}
    \textit{Each subdomain in the system has a set of nearest
      neighbors determined by the parallel adjacency graph of the
      input matrix. The subdomains are indexed by an integer.}}
  \label{fig:nearest_neighbor_comm}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{1.25}{
      \input{chapters/parallel_mc/binary_comm_tree.pdftex_t} }
    \caption{\textbf{Binary communication tree for coordinating the
        end of a parallel Neumann-Ulam solve.} \textit{Each child
        process reports to its parents how many histories completed
        within its domain. When the root process sums all complete
        histories, it forwards the stop signal to its children which
        in turn forward the message up the tree. The subdomains are
        indexed by an integer.}}
  \end{center}
  \label{fig:binary_comm_tree}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{0.75}{
      \input{chapters/parallel_mc/master_comm_tree.pdftex_t} }
    \caption{\textbf{Master/slave scheme for coordinating the end of a
        parallel Neumann-Ulam solve.} \textit{Each slave process
        reports to the master how many histories completed within its
        domain. When the master process sums all complete histories,
        it sends the stop signal to all slave processes. The
        subdomains are indexed by an integer.}}
  \end{center}
  \label{fig:master_comm_tree}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{0.6}{ \input{chapters/parallel_mc/msod_construction.pdftex_t} }
  \end{center}
  \caption{\textbf{MSOD construction for 4 sets with overlap.}
    \textit{The linear system is used to construct the Monte Carlo
      transport domain on the primary set. The Monte Carlo data
      structures are then broadcast amongst the blocks to all sets in
      the system.}}
  \label{fig:msod_construction}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \scalebox{0.6}{ \input{chapters/parallel_mc/msod_tally.pdftex_t} }
  \end{center}
  \caption{\textbf{MSOD tally reduction over blocks.} \textit{The
      tally computed independently in each set is reduced across the
      blocks to the primary set. The linear system vector is updated
      on the primary set. In this example the tally vector in the
      block containing upper-left subdomains is combined. Identical
      reduction operations occur for all other blocks
      simultaneously.}}
  \label{fig:msod_tally_reduction}
\end{figure}

@@

%%---------------------------------------------------------------------------%%
\section{Leadership-Class Parallel Scaling Studies\ }
\label{sec:leadership_scaling_studies}
For all cases, we explore scaling for just the Monte Carlo transport
kernel and MCSA as a whole iterative scheme. This will be necessary to
determine if the initial relaxation step required at each iteration is
a limiting factor in convergence. For each case, it will be
constructive to study the scaling performance as a function of
spectral radius - dictated by the cross sections and grid size to
diffusion length ratio. In addition, all of these studies should
hopefully have data from both the cluster and Titan and should all be
compared to Aztec GMRES and CG implementations with appropriate
preconditioning.

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_comm_parameters.pdf}
  \end{center}
  \caption{\textbf{Neumann Ulam parallel algorithm free parameter
      sensitivity.} \textit{The color scale represents the ratio of
      the runtime of a particular parameter variation with the runtime
      from the faster parameter variation. A value of 1.02 on the
      color scale corresponds to a runtime 2\% longer than the faster
      time. All runtimes were within 6.5\% of the fastest observed
      case.}}
  \label{fig:titan_comm_parameters}
\end{figure}

Compared to the reference work, there was also not a significant
change in the runtime over the values of parameters tested with all
runtimes reported within 6.5\% of the fastest case observed. For this
work, a history buffer size of 4096 and a buffer check period of 128
performed best for the intitial calculations. However, most values in
the range tested will give good performance. Brunner and Brantley
found that a value of 1024 was best for both parameters. For the
results observed here, these parameters were only 0.7\% slower than
the fastest case. It may be best practice to fix these parameters
internally at 1024 as both this work and the reference found these to
perform well several years apart on both different parallel machines
and for different transport sequences.

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_pure_strong.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition strong scaling absolute
      efficiency.}  \textit{Super-linear speed-up from memory
      thrashing in base case. MCLS is an order of magnitude slower
      arithmetically, causing the higher efficiency.}}
  \label{fig:titan_pure_strong}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_bvsm.pdf}
  \end{center}
  \caption{\textbf{Binary tree vs. master/slave communication scheme
      absolute efficiency.} \textit{Pure domain decomposition absolute
      strong scaling. The master/slave scheme will not scale beyond
      $O(1,000)$ cores due to load imbalance.}}
  \label{fig:titan_strong_bvsm}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_absolute.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling absolute
      efficiency.}}
  \label{fig:titan_weak_absolute}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_alg_eff.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling algorithmic
      efficiency.}}
  \label{fig:titan_weak_algorithmic}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_implementation.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling
      implementation efficiency.}}
  \label{fig:titan_weak_implementation}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_bvsm.pdf}
  \end{center}
  \caption{\textbf{Binary tree vs. master/slave communication scheme
      absolute efficiency.}  \textit{Pure domain decomposition
      absolute weak scaling. The master/slave scheme will not scale
      beyond $O(1,000)$ cores due to load imbalance.}}
  \label{fig:titan_weak_bvsm}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_time.pdf}
  \end{center}
  \caption{\textbf{Wall time in seconds to solution for each case.}}
  \label{fig:titan_strong_ms_time}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_eff.pdf}
  \end{center}
  \caption{\textbf{Absolute parallel efficiency relative to 16-core
      1-set base case.}}
  \label{fig:titan_strong_ms_eff}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_impeff.pdf}
  \end{center}
  \caption{\textbf{Strong scaling implementation efficiency relative
      to 16-core 1-set base case.}}
  \label{fig:titan_strong_ms_impeff}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_time.pdf}
  \end{center}
  \caption{\textbf{Wall time in seconds to solution for each case.}}
  \label{fig:titan_weak_ms_time}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_eff.pdf}
  \end{center}
  \caption{\textbf{Weak scaling absolute efficiency relative to
      16-core 1-set base case.}}
  \label{fig:titan_weak_ms_eff}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_impeff.pdf}
  \end{center}
  \caption{\textbf{Implementation parallel efficiency relative to 16-core
      1-set base case.}}
  \label{fig:titan_weak_ms_impeff}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_subdomain.pdf}
  \end{center}
  \caption{\textbf{Strong scaling absolute efficiency for pure domain
      decomposition.} \textit{The subdomain Monte Carlo solver
      improves the parallel efficiency and therefore time to solution
      by eliminating all parallel communication during the Monte Carlo
      solve.}}
  \label{fig:titan_strong_subdomain}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_subdomain.pdf}
  \end{center}
  \caption{\textbf{Weak scaling absolute efficiency for pure domain
      decomposition.} \textit{The subdomain Monte Carlo solver
      improves the parallel efficiency and therefore time to solution
      by eliminating all parallel communication during the Monte Carlo
      solve.}}
  \label{fig:titan_weak_subdomain}
\end{figure}

>>>>

Compared to the serial construction of the Neumann-Ulam decomposition,
data for all states that is required to be on process must be
collected in parallel. For the pure domain decomposition case, given
an input parallel decomposition for the Neumann-Ulam decomposition,
for all local states $m$ we require access to $P_{mn}$ and $W_{mn}$
(or $P^T_{mn}$ and $W^T_{mn}$) for the adjoint method for all possible
states $n$. Given by Figure~\ref{fig:diffusion_graph}, consider the
adjacency graph of a single state in the neutron diffusion matrix for
the model problem presented in Appendix~\ref{chap:diffusion_problem}.
\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{ \input{chapters/parallel_mc/stencil_graph.pdftex_t} }
  \end{center}
  \caption{\textbf{Neutron diffusion equation state adjacency graph.}
    \textit{Adjacent boundary states are collected by traversing the
      graph one step for all local boundary states and collecting the
      equations for those states that are not local. For a given
      transition in the random walk for the diffusion problem, a
      history at mesh point $(i,j)$ may transition to all adjacent
      mesh points including itself, corresponding to a global state
      transition from $m$ to $n$ in the linear system.}}
  \label{fig:diffusion_graph}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{ \input{chapters/parallel_mc/domain_to_domain.pdftex_t} }
  \end{center}
  \caption{\textbf{Nearest neighbor history communication sequence.}
    \textit{Each subdomain in the system has a set of nearest
      neighbors determined by the parallel adjacency graph of the
      input matrix. The subdomains are indexed by an integer.}}
  \label{fig:nearest_neighbor_comm}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.25}{
      \input{chapters/parallel_mc/binary_comm_tree.pdftex_t} }
    \caption{\textbf{Binary communication tree for coordinating the
        end of a parallel Neumann-Ulam solve.} \textit{Each child
        process reports to its parents how many histories completed
        within its domain. When the root process sums all complete
        histories, it forwards the stop signal to its children which
        in turn forward the message up the tree. The subdomains are
        indexed by an integer.}}
  \end{center}
  \label{fig:binary_comm_tree}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \scalebox{0.75}{
      \input{chapters/parallel_mc/master_comm_tree.pdftex_t} }
    \caption{\textbf{Master/slave scheme for coordinating the end of a
        parallel Neumann-Ulam solve.} \textit{Each slave process
        reports to the master how many histories completed within its
        domain. When the master process sums all complete histories,
        it sends the stop signal to all slave processes. The
        subdomains are indexed by an integer.}}
  \end{center}
  \label{fig:master_comm_tree}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \scalebox{0.6}{ \input{chapters/parallel_mc/msod_construction.pdftex_t} }
  \end{center}
  \caption{\textbf{MSOD construction for 4 sets with overlap.}
    \textit{The linear system is used to construct the Monte Carlo
      transport domain on the primary set. The Monte Carlo data
      structures are then broadcast amongst the blocks to all sets in
      the system.}}
  \label{fig:msod_construction}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \scalebox{0.6}{ \input{chapters/parallel_mc/msod_tally.pdftex_t} }
  \end{center}
  \caption{\textbf{MSOD tally reduction over blocks.} \textit{The
      tally computed independently in each set is reduced across the
      blocks to the primary set. The linear system vector is updated
      on the primary set. In this example the tally vector in the
      block containing upper-left subdomains is combined. Identical
      reduction operations occur for all other blocks
      simultaneously.}}
  \label{fig:msod_tally_reduction}
\end{figure}

@@

%%---------------------------------------------------------------------------%%
\section{Leadership-Class Parallel Scaling Studies\ }
\label{sec:leadership_scaling_studies}
Using the parallel performance metrics presented in the previous
section, we will next performance various scaling studies using an
implementation of the parallel MCSA algorithm to assess the quality of
the algorithm. For these scaling studies, large problems at high
levels of concurrency will be solved at a leadership-class computing
facility. In this section, we will outline the parallel machine on
which the calculations were performed and then perform a series a
studies that identify both strengths and weaknesses of the parallel
algorithm. Both pure domain decomposition and MSOD parallel
performance will be explored and compared to performance results for
the same problems using a set of production parallel Krylov solvers.

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_comm_parameters.pdf}
  \end{center}
  \caption{\textbf{Neumann Ulam parallel algorithm history buffer
      check period vs. history buffer size sensitivity.} \textit{The
      color scale represents the ratio of the runtime of a particular
      parameter variation with the runtime from the faster parameter
      variation. A value of 1.02 on the color scale corresponds to a
      runtime 2\% longer than the faster time. All runtimes were
      within 6.5\% of the fastest observed case.}}
  \label{fig:titan_comm_parameters}
\end{figure}

Compared to the reference work, there was also not a significant
change in the runtime over the values of parameters tested with all
runtimes reported within 6.5\% of the fastest case observed. For this
work, a history buffer size of 4096 and a buffer check period of 128
performed best for the intitial calculations. However, most values in
the range tested will give good performance. Brunner and Brantley
found that a value of 1024 was best for both parameters. For the
results observed here, these parameters were only 0.7\% slower than
the fastest case. It may be best practice to fix these parameters
internally at 1024 as both this work and the reference found these to
perform well several years apart on both different parallel machines
and for different transport sequences. Therefore, for the scaling
studies presented in the following sections, values of 1024 for both
the history buffer size and message check frequency will be used.

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_pure_strong.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition strong scaling absolute
      efficiency.}  \textit{Super-linear speed-up from memory
      thrashing in base case. MCLS is an order of magnitude slower
      arithmetically, causing the higher efficiency.}}
  \label{fig:titan_pure_strong}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_bvsm.pdf}
  \end{center}
  \caption{\textbf{Binary tree vs. master/slave communication scheme
      absolute efficiency.} \textit{Pure domain decomposition absolute
      strong scaling. The master/slave scheme will not scale beyond
      $O(1,000)$ cores due to load imbalance.}}
  \label{fig:titan_strong_bvsm}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_absolute.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling absolute
      efficiency.}}
  \label{fig:titan_weak_absolute}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_alg_eff.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling algorithmic
      efficiency.}}
  \label{fig:titan_weak_algorithmic}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_implementation.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling
      implementation efficiency.}}
  \label{fig:titan_weak_implementation}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_bvsm.pdf}
  \end{center}
  \caption{\textbf{Binary tree vs. master/slave communication scheme
      absolute efficiency.}  \textit{Pure domain decomposition
      absolute weak scaling. The master/slave scheme will not scale
      beyond $O(1,000)$ cores due to load imbalance.}}
  \label{fig:titan_weak_bvsm}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_time.pdf}
  \end{center}
  \caption{\textbf{Wall time in seconds to solution for each case.}}
  \label{fig:titan_strong_ms_time}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_eff.pdf}
  \end{center}
  \caption{\textbf{Absolute parallel efficiency relative to 16-core
      1-set base case.}}
  \label{fig:titan_strong_ms_eff}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_impeff.pdf}
  \end{center}
  \caption{\textbf{Strong scaling implementation efficiency relative
      to 16-core 1-set base case.}}
  \label{fig:titan_strong_ms_impeff}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_time.pdf}
  \end{center}
  \caption{\textbf{Wall time in seconds to solution for each case.}}
  \label{fig:titan_weak_ms_time}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_eff.pdf}
  \end{center}
  \caption{\textbf{Weak scaling absolute efficiency relative to
      16-core 1-set base case.}}
  \label{fig:titan_weak_ms_eff}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_impeff.pdf}
  \end{center}
  \caption{\textbf{Implementation parallel efficiency relative to 16-core
      1-set base case.}}
  \label{fig:titan_weak_ms_impeff}
\end{figure}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_subdomain.pdf}
  \end{center}
  \caption{\textbf{Strong scaling absolute efficiency for pure domain
      decomposition.} \textit{The subdomain Monte Carlo solver
      improves the parallel efficiency and therefore time to solution
      by eliminating all parallel communication during the Monte Carlo
      solve.}}
  \label{fig:titan_strong_subdomain}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_subdomain.pdf}
  \end{center}
  \caption{\textbf{Weak scaling absolute efficiency for pure domain
      decomposition.} \textit{The subdomain Monte Carlo solver
      improves the parallel efficiency and therefore time to solution
      by eliminating all parallel communication during the Monte Carlo
      solve.}}
  \label{fig:titan_weak_subdomain}
\end{figure}
